When are Fuzzy Extractors Possible?

### 2014-09-29 info-bounds/FuzzyExtPossibility.tex
wrote a lemma about log (beta-1)

We show this across different sketch values.  Consider some fixed sketch value $s$ and let $h_s \overset{def}= H_0(V | \sketch(V) = s)$.  That is, there are $2^{h_s}$ possible values for $V$ conditioned on $\sketch(V) = s$.  Furthermore, recall that 
\[
\log \expe_{s\in \sketch(V)} 2^{h_s} = \log \expe_{s\in \sketch(V)} 2^{H_0(V | \sketch(V) = s)}  = \tilde{H}_0(V | \sketch(V)) %\le   (\gamma-t+1)\log |\mathbb{F}|+1.
\]  
Conditioned on seeing the point $V$ there are $(|\mathbb{F}|-1)^{\gamma-1}/|\mathbb{F}|$ possible values for $Z$ with disjoint support outside of the sketched point.  Consider these possible values for $Z$ as bins to be filled with the $2^{h_{ss}}$ balls~(possible values of $V | \sketch(V)=ss$).  The average number of balls in each bin is maximized when the bins are filled equally.  That is, the average number of balls in each bin is bounded by the number of balls divided by the number of bins.  That is, 
\begin{align*}
\tilde{H}_0(V |Z  , \sketch(V) = ss) &\le \log \left(\frac{\text{\# bins}}{\text{\# balls}}\right)\\
&= \log \left(\frac{2^{h_{ss}}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \right)
\end{align*}
Then averaging over the possible values of $s$, we have the following:
\begin{align*}
\tilde{H}_0(V |Z , \sketch(V) ) &= \log \expe_{s\in \sketch(V)} 2^{\tilde{H}_0(V |  \sketch(V) =ss , (Z| \sketch(V) =ss) )}\\
&= \log\expe_{s\in \sketch(V)} \left(\frac{2^{h_s}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \right)\\
&= \log \frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\\
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \log \expe_{s\in \sketch(V)} 2^{h_s}\\
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \tilde{H}_0(V | \sketch(V))\\ 
&\le \log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + (\gamma-t+1)\log |\mathbb{F}|+1\\
&\le (\gamma-t+2)\log |\mathbb{F}| - (\gamma-1) \log (|\mathbb{F}|-1)+1\\
&\le (\gamma-t+2)\log |\mathbb{F}| - (\gamma-1) \log (\frac{|\mathbb{F}|}{2})+1\\
&\le (\gamma-t+2)\log |\mathbb{F}| - (\gamma-1) \log (|\mathbb{F}|) + (\gamma -1)+1\\
&\le (3-t)\log |\mathbb{F}| + (\gamma -1)+1\\
\end{align*}
This value is $\Theta(1)$ if $t\ge 3+ (\gamma-1)/\log | \mathbb{F}|$. \bnote{ I feel like this analysis can be tightened up.  Saying that $\log (|\mathbb{F}| -1 ) \ge \log (|\mathbb{F}|)-1$ seems like a big loss.}
 \end{proof}
\noindent \thref{thm:imposs sketch} follows using the change of quantifiers in \corref{cor:no fuzz for dist}.

>>>>

We show this across different sketch values.  Consider some fixed sketch value $s$ and let $h_s \overset{def}= H_0(V | \sketch(V) = s)$.  That is, there are $2^{h_s}$ possible values for $V$ conditioned on $\sketch(V) = s$.  Furthermore, recall that 
\[
\log \expe_{s\in \sketch(V)} 2^{h_s} = \log \expe_{s\in \sketch(V)} 2^{H_0(V | \sketch(V) = s)}  = \tilde{H}_0(V | \sketch(V)) %\le   (\gamma-t+1)\log |\mathbb{F}|+1.
\]  
Conditioned on seeing the point $V$ there are $(|\mathbb{F}|-1)^{\gamma-1}/|\mathbb{F}|$ possible values for $Z$ with disjoint support outside of the sketched point.  Consider these possible values for $Z$ as bins to be filled with the $2^{h_{ss}}$ balls~(possible values of $V | \sketch(V)=ss$).  The average number of balls in each bin is maximized when the bins are filled equally.  That is, the average number of balls in each bin is bounded by the number of balls divided by the number of bins.  That is, 
\begin{align*}
\tilde{H}_0(V |Z  , \sketch(V) = ss) &\le \log \left(\frac{\text{\# bins}}{\text{\# balls}}\right)\\
&= \log \left(\frac{2^{h_{ss}}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \right)
\end{align*}
Then averaging over the possible values of $s$, we have the following as long as $t\ge 4$:
\begin{align*}
\tilde{H}_0(V |Z , \sketch(V) ) &= \log \expe_{s\in \sketch(V)} 2^{\tilde{H}_0(V |  \sketch(V) =ss , (Z| \sketch(V) =ss) )}\\
&= \log\expe_{s\in \sketch(V)} \left(\frac{2^{h_s}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \right)\\
&= \log \frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\\
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \log \expe_{s\in \sketch(V)} 2^{h_s}\\
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \tilde{H}_0(V | \sketch(V))\\ 
&\le \log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + (\gamma-t+1)\log |\mathbb{F}|+1\\
&\le (\gamma-t+2)\log |\mathbb{F}| - (\gamma-1) \log (|\mathbb{F}|-1)+1\\
&< (\gamma-t+2)\log |\mathbb{F}| - (\gamma-2) \log |\mathbb{F}| +1 \ \ \ \ \mbox{(by \lemref{lem:log-minus-one})}\\
&\le (4-t)\log |\mathbb{F}| +1\\
& <1\,.
\end{align*}
\lnote{need for \lemref{lem:log-minus-one}:  $\gamma-1\le |F|$ and $|F|\ge 4$}
\lnote{change the any needed theorem statements it affects in light of the tightening}

\lnote{this lemma should probably live elsewhere since it is also needed for the FE case}
\begin{lemma}
\label{lem:log-minus-one}
For any real numbers $\alpha \leq \beta$ with $\beta \ge e+1$ (in particular, $b\ge 4$ suffices), the following holds:
$\alpha \log (\beta-1) > (\alpha-1)\log \beta$. 
\end{lemma}

\begin{proof}
Because $\beta-1$ is positive, and $1+x<e^x$ for positive $x$,
$$1+\frac{1}{\beta-1} < e^{\frac{1}{\beta -1}}\,.$$  Therefore, 
$$\left(1+\frac{1}{\beta-1}\right)^{\alpha-1} < e^{\frac{\alpha-1}{\beta-1}}\le e < \beta-1$$ (since $\alpha\le \beta$). Multiplying both sides by $(\beta-1)^{\alpha-1}$, we obtain
$$\beta^{\alpha-1} < (\beta-1)^\alpha\,.$$
Taking the logarithm of both sides, we get the desired result.
\end{proof}

 \end{proof}
\noindent \thref{thm:imposs sketch} follows using the change of quantifiers in \corref{cor:no fuzz for dist}.
